---
title: "Dectected toxins"
output:
  pdf_document: default
  html_document: default
date: "2026-01-08"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, message=FALSE}
library(here)
library(tidyverse)
library(broom)

load(here("data", "toxin_data.Rdata"))
```

Of the 34 toxins tested, 26 were found in at least one product. Of the 118 products tested, 104 had at least one detected toxin.

### Products with one or more toxins detected

```{r}
n <- main_data |> 
  select(ID) |> 
  distinct() |>
  nrow()

x <- main_data |> 
  summarize(n = n(), n_toxins = sum(amount > 0), .by = ID) |> 
  filter(n_toxins != 0) |> 
  nrow()

z <- epitools::binom.wilson(x, n) |> 
  mutate(across(c(proportion, lower, upper), ~ 100 * .x)) |> 
  rename(percent = proportion)

```

Products tested: `r z$n`

Products with one or more toxins detected: `r z$x`

Percent of products with one or more toxins detected: `r sprintf("%.1f",  z$percent)`; 95% CI = (`r sprintf("%.1f",  z$lower)`, `r sprintf("%.1f",  z$upper)`)

This is the distribution of the number of detected toxins: 14 products had no detected toxins, 17 had 1, etc.

### Distribution of number of detected toxins
```{r}
# distribution of number of toxins
number_toxins <- main_data |> 
  summarize(n = n(), n_toxins = sum(amount > 0), .by = ID)
# distribution of number of toxins
number_toxins |> 
  count(n_toxins)
```

Summary of the number of detected toxins:

```{r}
summary(number_toxins$n_toxins)
```

### Detected toxins by product type

Number of detected toxins by type of product, excluding "Miscellaneous"; e.g., of the 25 breakfast cereal products, 21 had at least one detected toxin.

```{r, message = FALSE}
by_food_type <- number_toxins |> 
  inner_join(products) |> 
  group_by(Food_type) |> 
  summarize(n = n(), any_toxin = sum(n_toxins > 0)) |> 
  mutate(no_toxin = n - any_toxin)
  # mutate(pct = 100 * any_toxin / n)

# Test for differences by type
# make contingency table as matrix
z1 <- by_food_type |> 
  filter(Food_type != "Miscellaneous") |> 
  select(any_toxin, no_toxin) |> 
  as.matrix(dimnames = list(by_food_type$Food_type, names(by_food_type)))
# row names not coming out for some reason; add manually
rownames(z1) <- by_food_type$Food_type[-3]
z1
```

A simulated chi-square test indicates that the proportions of products with at least one detected toxin are approximately equal across the six groups (P = 0.47).

```{r}
# chi-square approximation questionable, use simulated p-value
chisq.test(z1, simulate.p.value = TRUE)
```

This is the number of occurrences of each toxin in the 118 products tested.

```{r}
toxin_tally <- main_data |> 
  filter(amount > 0) |> 
  count(toxin) |> 
  arrange(-n)
print(toxin_tally, n = Inf)
```

These 8 toxins were not detected in any product.

```{r}
# toxins not occurring in any product
setdiff(unique(main_data$toxin), toxin_tally$toxin)
```

Most prevalent detected toxins by product group, excluding "Miscellaneous"; top five shown:

```{r}
food_type_list <- main_data |> 
  filter(amount > 0, Food_type != "Miscellaneous") |> 
  group_by(Food_type) |>
  count(toxin) |> 
  arrange(Food_type, -n) |> 
  group_split() 

walk(food_type_list, \(x) print(x, n = 5))
```

